---
title: "Creating_Biomass_Rasters"
author: "Lavran Pagano"
date: "1/17/2021"
output: html_document
---
#Part 1: Creating a raster with 10X10 meter spatial resolution based off of lidar derived covariets
Import the raster and car package
```{r}
library(raster)
library(car)
```

###Import canopy height raster and remove NAs
```{r}
rastblast <-raster("/Users/lavran_pagano/Downloads/Nan_as_0_masked_ham_chms/fixedmasked320320split1.tif")
rastblast[is.na(rastblast)]<- 0
```
###Plot the canopy height model
```{r}
plot(rastblast)
```


### Agregate a new raster with 10X10 meter resolution (133X133 pixels based off of the cell resolution) based off of 90th height percentiles
```{r}
radrast<-aggregate(rastblast, fact=133, fun=function(i, na.rm) quantile(i, probs=0.90, na.rm=na.rm))
```

###Plot the raster
```{r}
plot(radrast)
```


### Import training data
```{r}
Mastersheet<- read.csv('/Users/lavran_pagano/Desktop/Training_Data/Plot_Biomass_soft maplesTD.csv')
```

###Create model (in our case we are using a linear fixed effects model)
```{r}
latestmodel<-lm(sqrt(Biomass)~ sqrt(Height_90th) -1
                ,data=Mastersheet)
```
###Make sure all necesary assumptions of the model are met
```{r}
shapiro.test(residuals(latestmodel))
ncvTest(latestmodel)
```
###Summary staistics of the model we just made
```{r}
summary(latestmodel)
```
###Convert to a dataframe
```{r}
rastdataframe<-as.data.frame(radrast)
```
###Change name to indpendent variable name
```{r}
colnames(rastdataframe)[1]<-"Height_90th"
```
###Do necesary data transformations
```{r}
sqrtrastdataframe<-sqrt(rastdataframe)
```
###Predict biomass
```{r}
sqrtbiodata<-predict(latestmodel,rastdataframe)
```
###Do necesary data transformations part 2
```{r}
biodata<-sqrtbiodata^2
```
## Inspect the results
###Average Plot Biomass
```{r}
mean(biodata)
```
###Total Biomass of The Canopy Height Model
```{r}
sum(biodata)
```
# Part 2: Creating a Biomass Raster

###Get the cordinates of previous raster
```{r}
cord<-coordinates(radrast)
```
###Attach biomass values to a dataframe
```{r}
bioraster<-as.data.frame(cord)
bioraster$Biomass <-biodata
```
###Turn the dataframe into a raster
```{r}
bioraster<-rasterFromXYZ(bioraster) #Convert first two columns as lon-lat and third as biomass
```
###Plot the results
```{r}
plot(bioraster)
```


###Change the projection to the previous raster's
```{r}
rastcrs<-crs(rastblast)
crs(bioraster)<-rastcrs
```
###Write biomass raster to working directory
```{r}
setwd("/Users/lavran_pagano/Downloads/HamltonbiomassCHMs")
writeRaster(bioraster,"biomass320320split1.tif",overwrite=TRUE)
```

```
