---
title: "Creating_Plot_Polygons"
author: "Lavran Pagano"
date: "1/30/2021"
output: html_document
---
Import Packages
```{r}
library(raster)
library(rgdal)
```

Import shapefile of plot south west corner
```{r}
SWC<-shapefile('/Users/lavran_pagano/Downloads/Creekshead_Plots (1)/Creekshead_Plots.shp')
```



Get Lat-Long Cordinates from Plots
```{r}
long_lat<-data.frame(Long = SWC$ESRIGNSS_1 ,Lat = SWC$ESRIGNSS_L )#get lat and long cordinates
long_lat<-na.omit(long_lat) # remove NAs
```

Create a function to covert to UTM projection
```{r}
# convert to UTM FUnction
LongLatToUTM<-function(x,y,zone){
  xy <- data.frame(ID = 1:length(x), X = x, Y = y)
  coordinates(xy) <- c("X", "Y")
  proj4string(xy) <- CRS("+proj=longlat +datum=WGS84")  ## for example
  res <- spTransform(xy, CRS(paste("+proj=utm +zone=",zone," ellps=WGS84",sep='')))
  return(as.data.frame(res))
}
```

Apply UTM Function
```{r}
utm<-LongLatToUTM(long_lat$Long,long_lat$Lat,17)
```
Create a dataframe file with plot the UTM plot center cordinates
```{r}
utmEasting<-utm$X +5
utmNorthing<-utm$Y +5
utm_center <-data.frame(ID=utm$ID, Easting=utmEasting, Northing = utmNorthing)
```

Renane if not already in the appropriate format
```{r}
centroids <- utm_center
colnames(centroids)[1]<- "plot"
colnames(centroids)[2]<-"Easting"
colnames(centroids)[3]<-"Northing"
```
Get cordinates of plot corners

```{r}
radius <- 5
yPlus <- centroids$Northing+radius
xPlus <- centroids$Easting+radius
yMinus <- centroids$Northing-radius
xMinus <- centroids$Easting-radius
```

Calculate polygon coordinates for each plot centroid
```{r}
square=cbind(xMinus,yPlus,  # NW corner
             xPlus, yPlus,  # NE corner
             xPlus,yMinus,  # SE corner
             xMinus,yMinus, # SW corner
             xMinus,yPlus)  # NW corner again - close ploygon
```

Extract the plot ID information
```{r}
# Extract the plot ID information
ID=centroids$plot
```

Create spatial polygon dataframe 
```{r}
polys <- SpatialPolygons(mapply(function(poly, id) 
{
  xy <- matrix(poly, ncol=2, byrow=TRUE)
  Polygons(list(Polygon(xy)), ID=id)
}, 
split(square, row(square)), ID),
proj4string=CRS(as.character('+proj=utm +zone=17 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs')))
```

Create spatial polygon DataFrame and export as a shapefile
```{r}
## create spatial polygon DataFrame and export as a shapefile
polys.df <-SpatialPolygonsDataFrame(polys, data.frame(id=ID, row.names=ID))

# write to a polygon shapefile
writeOGR(polys.df, '.', 'CreeksheadPlotPolygon2', 'ESRI Shapefile',overwrite=TRUE)
```

