---
title: "Extracting_Lidar_Covariates"
author: "Lavran Pagano"
date: "1/30/2021"
output: html_document
---
Import Packages

```{r}
library(lidR)
library(raster)
library(purrr)
```

Import Polygon Plot Shapefile

```{r}
centShape <-shapefile("/Users/lavran_pagano/Downloads/CreeksheadPlotPolygon2.shp")
```

Import list of canopy height models and mosaic them into one large raster covering the study area

```{r}
#path to chm folder
mypath<- '/Users/lavran_pagano/Downloads/Lidar_Training_Data/Creeks_Head/Creeks_HeadCHMs'
#make a list of chm files
chmlist<-list.files(mypath,full.names = T, pattern = '.tif$') %>%  
map(raster)
#mosaic raster
mosaic<- do.call(merge,chmlist)
```

Make sure both raster and shapefile are the same projection. In our case we will be working in state plane

```{r}
#define projection (state plane)
statePlane<-"+proj=lcc +lat_0=41.5 +lon_0=-84.3666666666667 +lat_1=43.6666666666667
+lat_2=42.1 +x_0=3999999.999984 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=ft +no_defs"
#change chm projection
crs(mosaic) <- statePlane
#change shapefile projection
centShape<-spTransform(centShape,statePlane)
```

Create list of pixel values

```{r}
#create list of pixel values
pixelvalues<-raster::extract(mosaic,centShape,method= 'simple')
#Omit NAs so simplicity
pixelvalues <-lapply(pixelvalues,na.omit)
```

###Derive covariets

Height covariets
```{r}
#max hight
Height_Max<-lapply(pixelvalues,max)
Height_Max<-unlist(Height_Max)
#Min height
Height_Min<-lapply(pixelvalues,min)
Height_Min<-unlist(Height_Min)
#height 10th
percent10<-function(x){
  quantile(x, probs = c(0.1))
}
Height_10th<-lapply(pixelvalues,percent10)
Height_10th<-unlist(Height_10th)
#height 20th
percent20<-function(x){
  quantile(x, probs = c(0.2))
}
Height_20th<-lapply(pixelvalues,percent20)
Height_20th<-unlist(Height_20th)
#height 30th
percent30<-function(x){
  quantile(x, probs = c(0.3))
}
Height_30th<-lapply(pixelvalues,percent30)
Height_30th<-unlist(Height_30th)
#height 30th
percent50<-function(x){
  quantile(x, probs = c(0.5))
}
Height_50th<-lapply(pixelvalues,percent50)
Height_50th<-unlist(Height_50th)
#height 40th
percent40<-function(x){
  quantile(x, probs = c(0.4))
}
Height_40th<-lapply(pixelvalues,percent40)
Height_40th<-unlist(Height_40th)
#height 50th
percent50<-function(x){
  quantile(x, probs = c(0.5))
}
Height_50th<-lapply(pixelvalues,percent50)
Height_50th<-unlist(Height_50th)
#height 60th
percent60<-function(x){
  quantile(x, probs = c(0.6))
}
Height_60th<-lapply(pixelvalues,percent60)
Height_60th<-unlist(Height_60th)
#height 70
percent70<-function(x){
  quantile(x, probs = c(0.7))
}
Height_70th<-lapply(pixelvalues,percent70)
Height_70th<-unlist(Height_70th)
#height 80
percent80<-function(x){
  quantile(x, probs = c(0.8))
}
Height_80th<-lapply(pixelvalues,percent80)
Height_80th<-unlist(Height_80th)
#height 90
percent90<-function(x){
  quantile(x, probs = c(0.9))
}
Height_90th<-lapply(pixelvalues,percent90)
Height_90th<-unlist(Height_90th)
```

Crown gemoetric voulmne(cgv) covariets
cgv is calculated from the sumation of all pixel values in a plot with the minium height value subtracted from said pixel values. That number is then multiplyed by the width and length of a given pixel cell to convert to cubic meters.Our spatial resolution is 0.25ft which is equalivlent to 0.075m, Therfore given our spatail resolution cgv is calculated by sum(y-min(y)) 0.075m X 0.075m.

```{r}
#CGV FUll
CGV100<-function(y){
  sum(y-min(y))* 0.005625 #0.075m X 0.075m
}
cgv100th<-lapply(pixelvalues,CGV100)
cgv100th<-unlist(cgv100th)
#CGV50th
CGV50<-function(x, y){
  y<-ifelse(x< percent50(x),x,percent50(x))
  sum(y-min(x))* 0.005625
}
cgv50th<-lapply(pixelvalues,CGV50)
cgv50th<-unlist(cgv50th)
#CGV60th
CGV60<-function(x, y){
  y<-ifelse(x< percent60(x),x,percent60(x))
  sum(y-min(x))* 0.005625
}
cgv60th<-lapply(pixelvalues,CGV60)
cgv60th<-unlist(cgv60th)
#CGV70th
CGV70<-function(x, y){
  y<-ifelse(x< percent70(x),x,percent70(x))
  sum(y-min(x))* 0.005625
}
cgv70th<-lapply(pixelvalues,CGV70)
cgv70th<-unlist(cgv70th)
```

Create Training Data Data Frame
```{r}
#Make trainging data table
Training_data<-data.frame(Height_Max = Height_Max,
                          Height_90th = Height_90th,
                          Height_80th = Height_80th,
                          Height_70th= Height_70th,
                          Height_60th= Height_60th,
                          Height_50th=Height_50th,
                          Height_40th=Height_40th,
                          Height_30th=Height_30th,
                          Height_20th=Height_20th,
                          Height_10th=Height_10th,
                          Height_Min=Height_Min,
                          cgv100=cgv100th,
                          cgv70th= cgv70th,
                          cgv60th=cgv60th,
                          cgv50th=cgv50th)
print(Training_data)
```

Write Training Data to a csv
```{r}
write.csv(Training_data,'Creeks_Head_Covariates2.csv')
```

