---
title: "Model_Comparison"
author: "Lavran Pagano"
date: "3/2/2021"
output: html_document
---
###Import Packages
```{r}
# Import Packages
library(randomForest)
library(e1071)
library(MASS)
library(Metrics)
library(crossval)
```


###Import trainging data
```{r}
Mastersheet<- read.csv('/Users/lavran_pagano/Downloads/Lidar_Training_Data/Lidar_Training_Data.csv')
head(Mastersheet)
LogMastersheet <-log(Mastersheet[,-c(2:3)])
Mastersheet<-Mastersheet[,-c(2:3)]
```
#OLS
###Finding the best OLS model
```{r}
#OLS with all covariates
fit <- lm(Biomass~ Height_Max +Height_Min+cgv100+cgv70th+cgv60th+cgv50th+Height_10th+Height_20th+Height_30th+
            +Height_40th+Height_50th+Height_60th+Height_70th+Height_80th+Height_90th,data=Mastersheet)
```
###Stepwise AIC to determine best model
```{r}
step <- stepAIC(fit, direction="both")
step$anova # display results 
```


###OLS Model with the lowest AIC 
```{r}
AIC(lm<- lm(Biomass ~ cgv100 + cgv60th + Height_30th + Height_60th + Height_80th,data=Mastersheet))
summary(lm)
```

###10 Fold Cross Validation for OLS
```{r}
# set seed
set.seed(100)
# create a function for cross valdidated R^2 and RMSE
  lmfun <- function(train.x, train.y, test.x, test.y)
  {lm.fit <- lm(train.y ~ ., data=train.x)
    R2 <- cor(test.x, test.y) ^ 2
    predbio <- predict(lm.fit, test.x )
    RMSE <- rmse(predbio,test.y)
    stats<-data.frame(R2,RMSE)
    return(stats)
  }
# dataframe of lidar predictors from the model with the lowest AIC
x <- Mastersheet[c( "cgv100" , "cgv60th" , "Height_30th",  "Height_60th" ,"Height_80th")]
# vector of biomass
y <-Mastersheet$Biomass
# apply the crossval function
results <- crossval(lmfun,x,y, K=10)
#print the results
results$stat
```

#Power Law

###Finding the best Power Law model
```{r}
#Power Law all covariates (note any covariates with zeros were not considered)
fit <- lm(log(Biomass)~ log(Height_Max)+log(cgv100)+log(cgv70th)+log(cgv60th)+log(cgv50th)+log(Height_40th)+log(Height_50th)+log(Height_60th)+log(Height_70th)+log(Height_80th)+log(Height_90th),data=Mastersheet)
summary(fit)
```

###Stepwise AIC to determine best model
```{r}
step <- stepAIC(fit, direction="both")
step$anova # display results 
```

###Power law model with the lowest AIC
```{r}
AIC(powerlaw<-lm(log(Biomass) ~ log(Height_Max) + log(cgv100) + log(cgv70th) + log(cgv60th) + log(Height_80th) + log(Height_90th),data=Mastersheet))
summary(powerlaw)
```


###10 Fold Cross Validation for Powerlaw
```{r}
# set seed
set.seed(100)
# create a function for cross valdidated R^2 and RMSE
  plfun <- function(train.x, train.y, test.x, test.y)
  {pl.fit <- lm(train.y ~ ., data=train.x)
    R2 <- cor(test.x, test.y) ^ 2
    predbio <- predict(pl.fit, test.x )
    RMSE <- rmse(predbio,test.y)
    stats<-data.frame(R2,RMSE)
    return(stats)
  }
# dataframe of lidar predictors from the model with the lowest AIC
x <- LogMastersheet[c( "Height_Max", "cgv100" , "cgv70th" , "cgv60th", "Height_80th","Height_90th")]
# vector of biomass
y <-LogMastersheet$Biomass
# apply the crossval function
results <- crossval(plfun,x,y, K=10)
#print the results
results$stat
```

#Random forest
###  Random forest variable importance
```{r}
RandysForest<-randomForest(Biomass~ ., data= Mastersheet)
varImpPlot(RandysForest)
```

###10 Fold Cross Validation for Random Forest
```{r}
# set seed
set.seed(100)
# create a function for cross valdidated R^2 and RMSE
  rffun <- function(train.x, train.y, test.x, test.y)
  {rf.fit <- randomForest(train.y ~ ., data=train.x,type="eps-regression")
    R2 <- cor(test.x, test.y) ^ 2
    predbio <- predict(rf.fit, test.x )
    RMSE <- rmse(predbio,test.y)
    stats<-data.frame(R2,RMSE)
    return(stats)
  }
# dataframe of all lidar predictors from the model 
x <- Mastersheet[c( "Height_Max", "Height_Min","Height_10th","Height_90th","Height_20th")]
# vector of biomass
y <- Mastersheet$Biomass
# apply the crossval function
results <- crossval(rffun,x,y, K=10)
#print the results
results$stat
```
#SVR

###10 Fold Cross Validation for SVR
```{r}
# set seed
set.seed(100)
# create a function for cross valdidated R^2 and RMSE
  svrfun <- function(train.x, train.y, test.x, test.y)
  {svr.fit <- svm(train.y ~ ., data=train.x)
    R2 <- cor(test.x, test.y) ^ 2
    predbio <- predict(svr.fit, test.x )
    RMSE <- rmse(predbio,test.y)
    stats<-data.frame(R2,RMSE)
    return(stats)
  }
# dataframe of all lidar predictors from the model 
x <- Mastersheet[c( "Height_Max","Height_Min","Height_10th")]
# vector of biomass
y <- Mastersheet$Biomass
# apply the crossval function
results <- crossval(svrfun,x,y, K=10)
#print the results
results$stat
```

